
\chapter*{ABSTRACT}
\addcontentsline{toc}{chapter}{ABSTRACT}

Falls are among the leading causes of severe injuries in older adults and individuals requiring special monitoring. This thesis presents the design and implementation of a real-time fall detection and alert system, built on a fully integrated and distributed architecture.

The system comprises two primary detection components, each optimized for a specific monitoring scope. The first component is a wearable system, which employs inertial measurement units (IMUs) to track the user’s movements and posture, enabling large-scale and mobile monitoring. The second component is a computer vision system with fixed monitoring coverage, applying video analysis algorithms to observe critical areas in the environment, such as living rooms or bedrooms. When a fall event is detected by either component, the information is transmitted to a central processing platform.

The processing logic and control at the edge layer are executed on an ESP32 microcontroller, chosen for its balance of performance and cost-efficiency. Once a fall event is confirmed, alert data are encoded in JSON format and transmitted to the server using the MQTT protocol, ensuring minimal latency. This architecture enables the system to remain reliable even if one of the two components experiences a failure.

The system further integrates an automated telecommunication solution based on Asterisk PBX and a 4G-GPS module, which allows automatic initiation of emergency phone calls, transmission of SMS messages containing precise GPS coordinates, and sharing of on-site images through OTT messaging applications, thereby ensuring a fast and multi-channel emergency response.

The outcome is a modular, highly scalable architecture that can be readily extended with additional sensors or alert methods in the future. The system demonstrates a seamless integration of embedded IoT, automated telecommunications, and data analytics, contributing to improved quality of life and enhanced community safety.
